---
title: "Intermediate Presentation"
author: "Vy Lam, Graham Elliot"
date: "2023-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(vip)
library(tidymodels)
library(randomForest)
library(yardstick)
library(caret)
tidymodels_prefer()
conflicted::conflict_prefer("vi", "vip")

# Read in the data
movies <- read_csv("/Users/vy/Desktop/1-School/Fall 23/Projects in DS/movies.csv")

movies<-movies %>% 
  drop_na()
```


## Topic Description
We're currently working on exploring the realm of films and their production value.
Our goals are to figure out whether or not there is a set formula for a film's success. 
We've chosen a dataset from Kaggle. The dataset draws in around 200 plus movies each year
for the past four decades. 

Data Source: https://www.kaggle.com/datasets/danielgrijalvas/movies 

Our research question(s): 

1. Can we predict success of a film based on factors such as budget, runtime, and year?
2. How can we isolate individual variables best contribute to high revenue and positive ratings of a film?

While our expert is highly skeptical of data and the idea of it affecting a movie's performance, 
we also wanted to bring attention to how perhaps maybe quantifiable items aren't the only
influencers of a film's success.

### Data Wrangling
```{r}
movies <- movies %>%
  mutate(gross = factor(gross)) %>% #make sure outcome is factor
  mutate(across(where(is.character), as.factor))

movies <- movies %>% 
  subset(!grepl("e", gross))

movies$gross <- as.numeric(movies$gross)
```

### Randomizing 
```{r}
set.seed(456)
```

### Explanation

The following steps of random forest are all to pre-process the data & creating
workflows for the tree-making process.
```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = NULL, # size of random subset of variables
           trees = 1000, # Number of trees
           min_n = 2,
           importance = 'impurity') %>% 
  set_mode('regression') # change this for regression

# Recipe
data_rec <- recipe(gross ~., data = movies)

# Workflows
data_wf_mtry2 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 2)) %>%
  add_recipe(data_rec)

data_wf_mtry4 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 4)) %>%
  add_recipe(data_rec)

data_wf_mtry7 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 7)) %>%
  add_recipe(data_rec)

data_wf_mtry13 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 13)) %>%
  add_recipe(data_rec)
```

```{r}
set.seed(456) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry2 <- fit(data_wf_mtry2, data = movies)

set.seed(456)
data_fit_mtry4 <- fit(data_wf_mtry4, data = movies)

set.seed(456) 
data_fit_mtry7 <- fit(data_wf_mtry7, data = movies)

set.seed(456)
data_fit_mtry13 <- fit(data_wf_mtry13, data = movies)
```

```{r}
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          class = truth,
          model = model_label
      )
}
rf_OOB_output(data_fit_mtry2,'mtry2', movies %>% pull(gross))
```

```{r}
data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry2,'mtry2', movies %>% pull(gross)),
    rf_OOB_output(data_fit_mtry4,'mtry4', movies %>% pull(gross)),
    rf_OOB_output(data_fit_mtry7,'mtry7', movies %>% pull(gross)),
    rf_OOB_output(data_fit_mtry13,'mtry13', movies %>% pull(gross))
)
```

```{r}
data_fit_mtry2
data_fit_mtry4
data_fit_mtry7
data_fit_mtry13
```

```{r}
data_rf_OOB_output %>%
  group_by(model) %>%
  yardstick::rmse(.pred_class, class)
```

```{r}
ggplot(data_rf_OOB_output, aes(x = model, y = .pred_class - class)) +
  geom_boxplot() +
  labs(title = "Model RMSE Comparison", x = "mtry", y = "RMSE")
```
A model comparison for RMSE to see the average difference between values predicted by a model
and the actual values. This helps in figuring out which model will be the more accurate one to
view variable importance.

```{r}
model_output <-data_fit_mtry4 %>% 
    extract_fit_engine() 

model_output %>% 
    vip(num_features = 10) + theme_classic() #based on impurity

model_output %>% vip::vi() %>% head()
model_output %>% vip::vi() %>% tail()
```

The column graph above highlights the top variables in which have the most impact 
in the revenue of films. The top five categories are budget, year, score, company, and runtime. 

Why? 
- Random Forest is a tree-based model and does not require feature scaling. The algorithm requires 
partitioning, even if you apply Normalization then also yield the result would be the same. Random forest
yields better results than linear regression. However, we must not forego one methodology 
without some fact-checking. Below are single predictors graph which can help people understand 
how some variables hold more power than others.

We've explored some single predictors versus our response variable, gross (revenue).
And they pretty much sum up some of the results from the random forest model.

### Single Predictor Graphs

```{r}
ggplot(movies, aes(x = runtime, y = gross)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Scatter Plot with Line of Best Fit (Runtime)",
       x = "Runtime",
       y = "Gross")

ggplot(movies, aes(x = budget, y = gross)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter Plot with Line of Best Fit (Budget)",
       x = "Budget",
       y = "Gross")

```
Runtime vs Gross Graph:
 - Many movies tend to perform better in terms of revenue when the duration of the movie 
 runs between 80-150 minutes. This says something more about the culture around us than the dataset itself
 
Budget vs Gross Graph:
- Overall, most movies' performance based on budget plateaus once the budget reaches 100 million 
threshold. 
- Typically, a movie's budget shouldn't be this much.

#### Tableau Viz: 
https://public.tableau.com/app/profile/vy.lam/viz/ImaginingGlobalMovies/Sheet1?publish=yes 

Here, we are looking at other single variable predictors, country and year . 
Country is about the movie's country of origin. The goal is to see, aside from the variables
given by the random forest model, can we see success based on country and year of film's release?
Well, we tried to visualize the data using an interactive world map to see our data 
better as a scatterplot would not be suffice. We've decided highlight two years
and see how well they perform across all the countries that released a movie, and the data is in the dataset.
They are 2010 and 2020. In 2010, the country with the highest average revenue is the United States.
The goal is to get a decade analysis of maybe which country have higher averages
in revenue for a movie release. To our surprise, it's not the United States that had the highest
average in revenue for 2020. Through this exploration, we found that there might be some limitations 
within our datdasets and we might have to expand and join more data later on.
## Next Steps

Work plan: https://docs.google.com/document/d/1Yb6owP6v_fIm-QJlBFH_MZ9QL9_FiyhtHPV-e-SIZ0s/edit

## Contribution Summary

Graham contributed to:
 - Single predictor graphs in R & part of the analysis for them
 - Shiny App in R
 - Discussion of work plan & next steps
 - Reaching out to Leslie Mynt
 - Worked on putting the presentation together
 
Vy contributed to:
 - Running Random Forest Method & analysis, R analysis of a few graphs made in R
 - Creating interactive world map in Tableau 
 - Fleshing out work plan & discussing next steps
 - Worked on putting together the presentation
 - Revising graphs & compiling RMD for final submission
